TEMPERATURES AND METALLICITIES OF M DWARFS IN THE APOGEE SURVEY

### 2019-08-01 paper/draft.tex
clarify metallicity dist description

\begin{figure*}
\begin{center}
	\includegraphics[height=5cm]{figures/toomre.png}
	\includegraphics[height=5cm]{figures/met_hist.png}
	\caption{Toomre diagram showing the distribution of the \cannon\ metallicities (left). Dashed lines at $v_{tot}=70$\,km/s and 180\,km/s show roughly the separation of thin/thick/halo stars. Histogram (right) shows the metallicity distribution of thin and thick disk stars, with blue corresponding to $v_{tot}<70$\,km/s and green corresponding to $70<v_{tot}<180$\,km/s respectively. The average metallicity of the thin disk distribution is $\feh=0.00\pm0.17\,$dex, and the average metallicity of the thick disk distribution is $\feh=-0.14\pm0.19\,$dex.}
	\label{fig:toomre}
\end{center}
\end{figure*}

@@

As expected, the Toomre diagram in Figure \ref{fig:toomre} shows that higher metallicity sources in the sample are concentrated in low velocity space corresponding roughly to the thin disk population, while the thick disk population contains a higher concentration of lower metallicity sources. Separating the two populations into separate histograms (also shown in Figure \ref{fig:toomre}), we find that thick disk sources are more metal poor than thin disk sources, with the average metallicity of the thin disk stars being near-solar $\feh=0.00\pm0.17\,$dex, and the average metallicity of the thick disk stars being $\feh=-0.14\pm0.19\,$dex.
Metallicities of the two populations from ASPCAP show a similar distribution, with an average thin disk metallicity of $\mh=-0.16\pm0.16\,$dex and an average thick disk metallicity of $\mh=-0.23\pm0.17\,$dex.

@@

Given that physics-based spectral models of M dwarfs have issues, one of the possible
future values of the data-driven model shown here is that it is highly interpretable:
It contains within it first and second derivatives of the spectral expectation with
respect to the atmospheric parameters.
We show some of these derivatives in Figures \ref{fig:demo_teff}, \ref{fig:demo_feh}, and \ref{fig:west_derivative} and deliver relevant data in
Table~\ref{table:teff_derivative} and Table~\ref{table:feh_derivative}.
These Tables summarize spectral features in the APOGEE bandpass
that are found to be strong temperature and metallicity indicators.
In the long run, this is the primary value of data-driven models for astronomy:
To provide physical insights that drive physical understandings.
It is our hope that \thecannon, and models like it, will lead to
new and improved physical models which will, in turn, put \thecannon\ out of business. \\

>>>>

\begin{figure*}
\begin{center}
	\includegraphics[height=5cm]{figures/toomre.png}
	\includegraphics[height=5cm]{figures/met_hist.png}
	\caption{Toomre diagram showing the distribution of the \cannon\ metallicities (left). Dashed lines at $v_{tot}=70$\,km/s and 180\,km/s show roughly the separation of thin/thick/halo stars. Histogram (right) shows the metallicity distribution of thin and thick disk stars, with blue corresponding to $v_{tot}<70$\,km/s and green corresponding to $70<v_{tot}<180$\,km/s respectively. The average/standard deviation metallicity of the thin disk distribution is $\feh=0.00\pm0.17\,$dex, and the average/standard deviation metallicity of the thick disk distribution is $\feh=-0.14\pm0.19\,$dex.}
	\label{fig:toomre}
\end{center}
\end{figure*}

@@

As expected, the Toomre diagram in Figure \ref{fig:toomre} shows that higher metallicity sources in the sample are concentrated in low velocity space corresponding roughly to the thin disk population, while the thick disk population contains a higher concentration of lower metallicity sources. Separating the two populations into separate histograms (also shown in Figure \ref{fig:toomre}), we find that thick disk sources are more metal poor than thin disk sources, with the mean $\pm$ standard deviation being near-solar $\feh=0.00\pm0.17\,$dex for the thin disk distribution, and $\feh=-0.14\pm0.19\,$dex for the thick disk distribution.
Metallicities of the two populations from ASPCAP show a similar distribution, with the mean $\pm$ standard deviation being $\mh=-0.16\pm0.16\,$dex for thin disk sources and $\mh=-0.23\pm0.17\,$dex for thick disk sources.

@@

Given that physics-based spectral models of M dwarfs have issues, one of the possible
future values of the data-driven model shown here is that it is highly interpretable:
It contains within it first and second derivatives of the spectral expectation with
respect to the atmospheric parameters.
We show some of these derivatives in Figures \ref{fig:demo_teff}, \ref{fig:demo_feh}, and \ref{fig:west_derivative} and deliver relevant data in
Table~\ref{table:teff_derivative} and Table~\ref{table:feh_derivative}.
These Tables summarize spectral features in the APOGEE bandpass
that are found to be strong temperature and metallicity indicators.
In the long run, this is the primary value of data-driven models for astronomy:
To provide physical insights that drive physical understandings.
It is our hope that \thecannon, and models like it, will lead to
new and improved physical models which will, in turn, put \thecannon\ out of business.
